---
title: "How well do you excercise?: Predicting your excercise using accelerometer's data"
author: "Gabriela Rodriguez"
date: "January 17th, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants to predict how well are they doing a barbell lift. 

## Libraries and loading data

These are the libraries that will be used in this work. 

We will download the training and testing data sets

```{r}
library(caret)
library(randomForest)
library(rpart)
library(readr)

pml_training <- read.csv("C:/Users/Gaby Rodriguez/Dropbox/Machine Learning Final project/pml-training.csv", na.strings=c("NA","#DIV/0!", ""))

pml_testing <- read.csv("C:/Users/Gaby Rodriguez/Dropbox/Machine Learning Final project/pml-testing.csv", na.strings=c("NA","#DIV/0!", ""))




```


## Cleaning data

Let's do some cleaning to our data sets, first we are going to remove all NA's, and then the first seven columns that are not needed for the aims of this project.

```{r}
trainingset <- pml_training[,colSums(is.na(pml_training)) == 0]
testingset <- pml_testing[,colSums(is.na(pml_testing)) == 0]

trainingset <- trainingset[, -c(1:7)]
testingset <- testingset[,-c(1:7)]

```


## Partition rows into training and crossvalidation
For reproducibility we are going to set our seed in 123. Next we are going to randomly use 75% of our training data set to train the model, the rest (25%) will be used to cross validate our model.

```{r}
set.seed(123)
inTrain <- createDataPartition(y=trainingset$classe, p=0.75, list=FALSE)
training <- trainingset[inTrain,]
crossv <- trainingset[-inTrain,]

```

## Models

Let's now try two models, first, random forest and second a decision tree

```{r}
mod1 <-randomForest(classe ~.,data=training,method="class")
mod2 <-rpart(classe ~.,data=training,method="class")


pred1 <-predict(mod1,crossv)
pred2 <-predict(mod2,crossv, type="class")

```

##Confusion matrices

```{r}
print("Model 1")
confusionMatrix(pred1,crossv$classe)
print("Model 2")
confusionMatrix(pred2,crossv$classe)

```

Comparing both matrices it is evident that the random forest algorithm has better accuracy, sensitivity and specificity 
 
## Submission

```{r}
prediction<- predict(mod1,testingset,type="class")
prediction

pml_files =function (x) {
  n= length(x)
  for(i in 1:n ){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename, quote = FALSE, row.names= FALSE, col.names=FALSE)
  }
}

pml_files(prediction)
```

##References


Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 
http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har

